A Realistic Guide to Making Data Available Alongside Code to Improve Reproducibility.

### 2019-12-09 paper.Rmd
tidy up authorship addresses; add sentence on open data commons open database license; expand on JSON vs XML; expand on dryad

---
title: A Realistic Guide to Making Data Available Alongside Code to Improve Reproducibility.
author: Nicholas J Tierney^1,2,*^, Karthik Ram^3,*^
output:
  bookdown::pdf_document2:
    toc: no
    includes:
      in_header: preamble.tex
bibliography: references.bib
header-includes: 
  - \linespread{2} 
  - \usepackage[utf8]{inputenc}
---

1 = Monash University, Department of Econometrics and Business Statistics
2 = Australia Centre of Excellence for Mathematical and Statistical Frontiers
3 = University of California, Berkeley
* = Both authors contributed equally to the work

@@

> "Data! data! data!" he cried impatiently. "I can't make bricks without clay." - Sherlock Holmes (PBS Series, 2010-2017)

@@

Other licenses or notices to be aware of are **copyrighted data**, and **data embargoes**. If you are working with data **already copyrighted**, (for example under CC BY or CC0), you must give follow appropriate guidelines for giving credit. Data may also be under **embargo**. This means data cannot be shared more widely until a specific release time. If sharing data under an embargo, include detailed information on the embargo requirements in: the README, and in separate correspondence with those who receive the data.

@@

In contrasst to a CSV data dictionary, JSON-LD has a defined, nested structure, which means more can be communicated efficiently, while still maintaining readability, and avoiding extra writing and repetition that comes with a plain CSV where everything is defined from scratch.

Another rich format is XML, the eXtensible Markup Language. This was an early iteration on the idea of a plain text format that was both human and machine readable. It is very powerful - it powers the entire Microsoft Office suite, but for metadata purposes it is not as human readable. However, one example of its usage is in EML, the Ecological Metadata Language. XML and JSON can be converted into each others respective formats

Expand on:

- The holy grail of the semantic web: JSON-LD

- readers just been advised to make a:
  - a csv data dictionary,
  - put that in the README
  - then make a JSON file. 
Is there a way to discuss strategies on harmonising all these formats?

- we could map out how these things link together (e.g., you can turn JSON into xml with package ...)
- make a table that describes how to do something in R, or Python, or Julia

@@

If the dataset you are shipping has a research application, the most relevant home for it would be the research data repositories **zenodo** or **dryad**. Zenodo, Launched in 2013 in a joint collaboration between openAIRE and CERN, provides a free, archival location for any researcher to depositor their datasets. The only limits to file sizes are 50gb for individual files, which is generous enough to accommodate a large number of use cases. Zenodo is able to accommodate larger file sizes upon request. Dryad performs data curation

```{r create-draft-table, }
tibble::tribble(
  ~Organisation,              ~Pros, ~Cons,
  "FigShare",                 "",     "",
  "Dryad",                    "",     "",
  "Zenodo",                   "",     "",
  "Institutional Repository", "",     "",
) %>% 
  knitr::kable(caption = "Table Comparing common data submission platforms")
```

@@

- Mile McBain
- Anna Krystalli
- Daniella Lowenberg

>>>>

---
title: A Realistic Guide to Making Data Available Alongside Code to Improve Reproducibility.
author: Nicholas J Tierney\* (1,2) & Karthik Ram\* (3)
output:
  bookdown::pdf_document2:
    toc: no
    includes:
      in_header: preamble.tex
bibliography: references.bib
header-includes: 
  - \linespread{2} 
  - \usepackage[utf8]{inputenc}
---

1 = Monash University, Department of Econometrics and Business Statistics

2 = Australia Centre of Excellence for Mathematical and Statistical Frontiers

3 = University of California, Berkeley

\* = Both authors contributed equally to the work

@@

> "Data! data! data!" he cried impatiently. "I can't make bricks without clay." - Sherlock Holmes (The Adventure of the Copper Beeches by Sir Arthur Conan Doyle)

@@

Other licenses or notices to be aware of are **copyrighted data**, and **data embargoes**. If you are working with data **already copyrighted**, (for example under CC BY or CC0), you must give follow appropriate guidelines for giving credit. Data may also be under **embargo**. This means data cannot be shared more widely until a specific release time. If sharing data under an embargo, include detailed information on the embargo requirements in: the README, and in separate correspondence with those who receive the data. Databases may also be licensed with the Open Data Commons Open Database License: https://opendatacommons.org/licenses/odbl/, which provides provisions for sharing, creating, and adapting, provided that work is attributed, shared, and kept open.

@@

In contrast to a CSV data dictionary, JSON-LD has a defined, nested structure, which means more can be communicated efficiently, while still maintaining readability, and avoiding extra writing and repetition that comes with a plain CSV where everything is defined from scratch.

Another rich format is XML, the e**X**tensible **M**arkup **L**anguage. This was an early iteration on the idea of a plain text format that was both human and machine readable. It is very powerful - it powers the entire Microsoft Office suite, but for metadata purposes it is not as human readable. XML is used is in EML, the Ecological Metadata Language. XML differs to JSON in that JSON is shorter, and quicker and easier to read and write. JSON can also be parsed with standard JavaScript, whereas XML must be parsed by a special XML parser. JSON has key-value pairs of information, whereas XML represents data in a tree structure. XML and JSON can be converted into each others respective formats, for example with the `json` or `xml2` R packages.

@@

If the dataset you are shipping has a research application, the most relevant home for it would be the research data repositories **zenodo** or **dryad**. Zenodo, Launched in 2013 in a joint collaboration between openAIRE and CERN, provides a free, archival location for any researcher to depositor their datasets. The only limits to file sizes are 50gb for individual files, which is generous enough to accommodate a large number of use cases. Zenodo is able to accommodate larger file sizes upon request. Dryad will take data from any field of research, and perform human quality control and assistance  of the data, with the ability to link data with a journal publication, but requires a data publishing fee.

```{r create-draft-table}
tibble::tribble(
  ~Organisation,              ~Pros, ~Cons,
  "FigShare",                 "",     "",
  "Dryad",                    "",     "",
  "Zenodo",                   "",     "",
  "Institutional Repository", "",     "",
) %>% 
  knitr::kable(caption = "Comparison of common data submission platforms.")
```

@@

- Mile McBain
- Anna Krystalli
- Daniella Lowenberg
- ACEMS International Mobility Programme
